S1: Extended Derivations for `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-28 ltx/PLoS/LaTeX_Manuscript.tex
added to acknowledgements
commiting to merge

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
With recent advances in medical imaging techniques and computing power we are entering a new age of neuroscientific research, where the amount of patient specific information is becoming incredibly detailed. In order to make the best use of this information, patient specific mathematical models should be formed. This is particularly poignant where considering poorly understood complex systems such as large scale dynamics of the human brain. In terms of brain dynamics, there currently exists a library of neural field models that provide a theoretical explanation of various phenomena associated with various cognitive states and disease. In this paper we have created a platform for creating a patient specific neural field model. Within this platform, we propose rules for designing experiments and derive mathematical formula to create such a model.

@@

   	\caption{{\bf Mexican-hat connectivity kernel, $w(\mathbf{r},\mathbf{r'})$}. The function $w(\cdot)$ is rotationally symmetric (isotropic) about zero, hence a cross-section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.}

@@

The frequency analysis and model selection tools presented in Section~\ref{SpectralAnalysisSection} specify the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to design electrophysiological recording systems to avoid spatial aliasing. In addition, the frequency analysis highlights the relationship between the spatial width of the sensors and the information in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009},

@@

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}
Dean Freestone would like to thank Dr. Mark van Rossum for hosting him whilst visiting the University of Edinburgh where this work was undertaken.

>>>>

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
With recent advances in medical imaging techniques and computing power we are entering a new age of neuroscientific research, where patient-specific information is becoming incredibly detailed. In order to make the best use of this information, patient-specific mathematical models should be formed. This is particularly poignant where considering poorly understood complex systems such as large-scale dynamics of the human brain. There currently exists a library of mathematical models describing large-scale brain dynamics, providing a theoretical explanation of various phenomena associated with various cognitive states and disease. In this paper, we have created a framework for creating a patient specific mathematical model, forming a bridge between theory and electrophysiological data. Within this framework, we propose rules for designing experiments and we derive the mathematical formula to create the model. To demonstrate the utility of this framework, we provide an example using synthetic data that shows a comparison between true and estimated model characteristics. The ability to create patient-specific models of large-scale brain dynamics has the potential to revolutionise the treatment of diseases such as epilepsy, where techniques from engineering may be used to prevent abnormal neural oscillations.

@@

   	\caption{{\bf Mexican-hat connectivity kernel.} The function $w(\cdot)$ is rotationally symmetric (isotropic) about zero, hence a cross-section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.}

@@

The frequency analysis and model selection tools presented in Section~\ref{SpectralAnalysisSection} specify the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to design electrophysiological recording systems to avoid spatial aliasing. In addition, the frequency analysis highlights the relationship between the spatial width of the sensors and the information in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009}.

@@

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}
Dean Freestone would like to thank Dr. Mark van Rossum for hosting him whilst visiting the University of Edinburgh where this work was undertaken. Authors would also like to thank Andrew Hills for useful discussions and help with some code and the generation of figures.

### 2010-06-28 ltx/PLoS/SupportingInformation.tex
editing discussion, including mikes edits

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi
\title{S1: Extended Derivations for `A Data Driven Framework for Patient-Specific Neural Field Modelling'}
\author{Dean R. Freestone$^{1,2,3}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{5,\ast}$,\\
Kenneth Scerri$^{6,\ast}$,
David B. Grayden$^{1,\ast}$,
Visakan Kadirkamanathan$^{4}$}

@@

@@

\maketitle

\begin{flushleft}

\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
\\
$\ast$ E-mail: dfreestone@bionicear.org
\end{flushleft}
\renewcommand{\theequation}{S1.\arabic{equation}}
\section*{Time Discretisation of Neural Field Model}\label{Time Discretization} To form the IDE neural field model a 
time discretisation must be performed. To do this we used a one-step Euler method, where equation~7 of the main text can be approximated by 
\begin{equation}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} =-\zeta v\left( \mathbf{r},t \right) + \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~S1.1 we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

>>>>

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi
% \title{S1: Extended Derivations for `A Data Driven Framework for Patient-Specific Neural Field Modelling'}
% \author{Dean R. Freestone$^{1,2,3,\ast}$, 
% Parham Aram$^{4}$, 
% Michael Dewar$^{5}$,\\
% Kenneth Scerri$^{6}$,
% David B. Grayden$^{1}$,
% Visakan Kadirkamanathan$^{4}$}

@@

	\begin{flushleft}
	{\Large
	\textbf{S1: Extended Derivations for `A Data Driven Framework for Patient-Specific Neural Field Modelling'}
	}
	% Insert Author names, affiliations and corresponding author email.
	\\
	Dean R. Freestone$^{1,2,3,\ast}$, 
	Parham Aram$^{4}$, 
	Michael Dewar$^{5}$,
	Kenneth Scerri$^{6}$,
	David B. Grayden$^{1}$,
	Visakan Kadirkamanathan$^{4}$
	\\
	\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
	\\
	\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
	\\
	\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
	\\
	\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
	\\
	\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
	\\
	\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
	\\
	$\ast$ E-mail: dfreestone@bionicear.org
	\end{flushleft}

@@

% \maketitle
% 
% \begin{flushleft}
% 
% \bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
% \\
% \bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
% \\
% \bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
% \\
% \bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
% \\
% \bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
% \\
% \bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
% \\
% $\ast$ E-mail: dfreestone@bionicear.org
% \end{flushleft}
\renewcommand{\theequation}{S1.\arabic{equation}}
\section*{Time Discretisation of Neural Field Model}\label{Time Discretization} To form the IDE neural field model a 
time discretisation must be performed. To do this we used a one-step Euler method, where equation~7 of the main text can be approximated by 
\begin{equation}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} =-\zeta v\left( \mathbf{r},t \right) + \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~S1.1 we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.
